---
title: "Internet Users"
author: "Andrew Ehsaei"
date: "February 16, 2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Packages
library(dplyr)
library(ggplot2)
library(knitr)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(GGally)
library(scales)
library(memisc)
library(tidyr)
library(gridExtra)
```
## Abstract
In this project I'll be creating an interactive data visualization using
HTML, CSS, JavaScript, and D3. The dataset I am exploring contains the numbers of 
Internet Users in each country over time.

## Introduction
First, I will download the dataset from the World Bank website
( http://data.worldbank.org/indicator/IT.NET.USER.P2 ). I'll start to explore the 
data using R and create some basic plots using ggplot to look for interesting 
trends. The README.md will contain details on creating the visualization.

## Import the Data

```{r echo=TRUE, load_data}
# Import the dataset into a dataframe variable
mydata <- read.csv('/Users/aehsaei/Documents/Udacity/Project_6/InternetUsersData.csv', row.names=NULL)
```

## Tidy Data
```{r echo=TRUE, tidy_the_data}
# Drop the Indicator.Name column (not needed)
mydata <- mydata[ , !(names(mydata) %in% c("Country.Code", "Indicator.Name"))]

# Change the 'Year' columns to row values
mydata <- melt(mydata, id=(c("Country.Name", "Indicator.Code")))

# Add column names to these new aggregated columns
colnames(mydata) <- c("Country.Name", "Indicator.Code", "Year", "Users")

# Remove the leading 'X' in front of the years
mydata["Year"] <- lapply(mydata["Year"], function(x) { gsub("X", "", x) })

# Change all 'NA' values to 0 in the Users column
mydata$Users[which(is.na(mydata$Users))] <- 0

# Print the transformed dataframe
str(mydata, strict.width = "cut", indent.str = "  " )
```

```{r echo=TRUE, explore_data}
# Find internet history of USA
subset(mydata, mydata$Country.Name == "United States" & mydata$Users > 0)

# Find first year with non-zero internet user data
subset(mydata, mydata$Users > 0)["Year"][1,]

# Remove all data before first year with non-zero values
mydata <- subset(mydata, mydata$Year >= 1990)

# Find top 10 countries with highest 'Users' in 2015
top_10_countries <- subset(mydata, mydata$Users > 0 & mydata$Year == 2015) %>%
  arrange(desc(Users)) %>%
  head(10)

top_10_countries

# Find lowest 10 countries with highest 'Users' in 2015
bottom_10_countries <- subset(mydata, mydata$Users > 0 & mydata$Year == 2015) %>%
  arrange(desc(Users)) %>%
  tail(10)

bottom_10_countries
```

```{r echo=TRUE, plot_data}

# Plot top 10
ggplot(data = subset(mydata, 
                     mydata$Country.Name %in% top_10_countries$Country.Name), 
       aes(x=Year, y=Users, group=Country.Name, colour=Country.Name)) +
  geom_line() +
  ggtitle("Countries With The Most Internet Users") +
  labs(x="", y="Users (Per 100 People)") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1))

# Plot bottom 10
ggplot(data = subset(mydata, 
                     mydata$Country.Name %in% bottom_10_countries$Country.Name), 
       aes(x=Year, y=Users, group=Country.Name, colour=Country.Name)) +
  geom_line() +
  ggtitle("Countries With The Least Internet Users") +
  labs(x="", y="Users (Per 100 People)") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1))

# Plot all
ggplot(data = mydata, aes(x=Year, y=Users, group=Country.Name)) +
  geom_line() +
  ggtitle("Internet Users (All Countries)") +
  labs(x="", y="Users (Per 100 People)") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1))
```

## Output Tidied Data
```{r echo=TRUE, export_the_data}
write.csv(file='/Users/aehsaei/Documents/Udacity/Project_6/InternetUsersDataFixed.csv', x=mydata)
```